---
title: "Time Spent on Athletics and Academics"
subtitle: "SAT 231: Calendar Query"
author: "Sofia Guttmann"
date: today
format: pdf
linestretch: 1.15
highlight-style: arrow
---

```{r}
#| label: setup
#| include: false

# set code chunk option defaults
knitr::opts_chunk$set(
  # display code as types
  tidy = FALSE, 
  # slightly smaller code font
  size = "small",
  # do not display messages in PDF
  message = FALSE,
  # set default figure width and height
  fig.width = 5, fig.height = 3) 

# improve digit and NA display 
options(scipen = 1, knitr.kable.NA = '')

# load packages
library(tidyverse)
library(lubridate)
library(ical)
# do you need more packages?
```

<!--
HOW TO USE THIS TEMPLATE:

The template is designed to help you structure your report. All work herein should be your own. Thus, you should write your own code comments, set your own chunk options, etc. You should clean up code comments that I've left so they are useful for you. 

You can leave the markdown comments for guidance in writing your report since they will not appear in the knitted pdf. If you like, you can remove them. 

Please rename/remove the headings as you see fit. 
You may add subheadings as desired to break up the report. Practice your Markdown formatting to make a professional looking report. 

REMINDERS ABOUT FILE ORGANIZATION AND GITHUB:
You should commit and push all relevant files to GitHub. Someone else (me, in particular) should be able to download the *calendar-query* folder from GitHub and run your .Qmd file without changing anything (meaning file paths are not specific to your computer and all files that are read in are provided.)

Your final *calendar-query* project folder should be organized such that file names and any subfolder names are informative, making relevant content easy to find for someone with little or no context of the project (i.e., if someone saw a file called "calendar_report_template", they would likely assume it was a template and not a finalized report). 
-->

# Introduction 
1. Time spent in class versus doing homework/studying
This question is important because for every hour spent in class, I want to be spending 2.5-3 hours doing supplemental studying, reading, and assignments. I want to make sure my time is being spent and allocated properly.

2. Time spent on academics versus athletics
This question is important because now that I am in the middle of lacrosse season, my time is very valuable.  I want to maintain my time for studies because my time for athletics is not very flexible.  I want to see how many hours in the day are left over that I can use for my studies.
<!--
Describe the questions of interest: what are they and why are they important or useful to you?
-->

# Data collection 

<!--
Briefly describe your data collection process, including how you defined variables of interest (include the levels/categories of categorical variables and the units for quantitative variables).
--> 
I took note of every time I would study, do an assignment, go to class, go to a workout, go to practice, or go to office hours over the course of 14 days.  The variables are hours (quantitative) and day of the week (qualitative).  

<!--
The code below is provided as a starting point for importing an `ics` file into R as a dataframe, doing some initial wrangling, and adding up the time spent on each activity. You are not required to use this code, but I hope it is useful as you get started. Feel free to delete or modify it as you see fit. 

If you are using the code, be sure to update the capitalized parts with the relevant path to where your `ics` file is located *relative to where this file is located in the repo*, and set `eval = TRUE`.

NOTE: All of your wrangling should be in this file - preferably well-organized and in one place if possible. You may use multiple chunks if it makes it easier to tell your narrative. This is demo-ed below, where the data is wrangled first for one visualization, then could be used for the second, etc. 
-->


```{r}
#| label: gcal-starter-code
#| eval: false #be sure to set this to true and/or just remove
 
# Load required libraries
library(ical)
library(tidyverse)
library(lubridate)

# Data import using the correct Google Calendar URL
cal_import <- ical_parse_df("https://calendar.google.com/calendar/ical/c_e5fb39631c03a80b1dcb7ed49c5f211177a58e144adddca1af0ac5220032aa7c%40group.calendar.google.com/public/basic.ics")

# Data wrangling
mycal <- 
  cal_import %>% 
  # Google Calendar event names are in a variable called "summary";
  # "activity" is a more relevant/informative variable name.
  rename(activity = summary) %>% 
  mutate(
    # Specify time zone (defaults to UTC otherwise)
    across(c(start, end), 
           .fns = with_tz, 
           tzone = "America/New_York"),
    # Compute duration of each activity in hours
    duration_hours = interval(start, end) / hours(1),
    # Examples of getting components of dates/times 
    # Note: 
    # i. these could be based on either start datetime or end datetime
    # ii. you do NOT need all of these!! so only use what you need
    date = date(start),
    year = year(start),
    month_number = month(start),
    month_label = month(start, 
                        label = TRUE, 
                        abbr = FALSE),
    weekday_number = wday(start),
    weekday_label = wday(start, 
                         label = TRUE, 
                         abbr = FALSE),
    hour = hour(start),
    time = hour(start) + minute(start)/60,
    # Convert text to lowercase and remove repeated or leading/trailing 
    # spaces to help clean up inconsistent formatting.
    across(c(activity, description), 
           .fns = str_to_lower),
    across(c(activity, description), 
           .fns = str_squish)
  ) %>% 
  # The first Google Calendar entry is always an empty 1969 event
  filter(year != 1969)


```


<!--
Briefly describe and justify your plans for analysis (i.e., visualizations and summary table). Which graphs and corresponding visual cues do you plan to use? Which summary measures will you use in your table? (This will likely be a less detailed version of what you submitted for your proposal.) The idea is to let your reader know what is coming before they get to it. 
--> 

# Results 
```{r eval=FALSE}
# Group by weekday_label and calculate total time spent in each activity
mycal_summary <- mycal_pivot %>%
  group_by(weekday_label) %>%
  summarise(
    Class = sum(class, na.rm = TRUE),         # Corrected from 'Class' to 'class'
    Practice = sum(practice, na.rm = TRUE),
    Academics = sum(Academics, na.rm = TRUE)
  ) %>%
  ungroup()

# Arrange the weekdays in the correct order
weekdays_order <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
mycal_summary$weekday_label <- factor(mycal_summary$weekday_label, levels = weekdays_order)

# Display the table
mycal_summary

```
One table I will use will have the day of the week in one column, time spent in athletics in the next column, time spent in the classroom in the next column, and time spent doing homework/studying in the next column.

```{r eval=FALSE}
library(ggplot2)

# Filter data for 'class' activity
class_data <- mycal_filtered %>%
  filter(activity == "class")

# Create a scatter plot
ggplot(class_data, aes(x = duration_hours, y = duration_hours)) +
  geom_point() +  # Add points
  labs(x = "Time spent in Academics", y = "Time spent in Class") +  # Label axes
  ggtitle("Time Spent in Class vs. Academics") +  # Add title
  theme_minimal()  # Apply a minimal theme


```
I will use a scatter plot to display the first question.  The y axis will be time spent in class and the x axis will be time spent doing homework/studying.


```{r eval=FALSE}
library(ggplot2)

# Filter data for "Academics" and "Practice" activities
filtered_activities <- mycal_filtered %>%
  filter(activity %in% c("Academics", "practice"))

# Create a bar graph
ggplot(filtered_activities, aes(x = weekday_label, y = duration_hours, fill = activity)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Day of the Week", y = "Hours") +
  ggtitle("Hours Spent on Academics and Practice by Day of the Week") +
  scale_fill_manual(values = c("Academics" = "blue", "practice" = "red")) +
  theme_minimal()


```
I will use bar graphs to display the second question.  The x axis will be the different days of the week with two bars in each day.  One of academics and one for athletics.  The y axis will represent time.


# Conclusions 

<!--
With your initial questions in mind, briefly summarize what you learned about how you spent your time. What are the big takeaways for you about how you spend your time? If relevant, how does this affect how you might spend your time going forward?
-->

Through this project, although I am not sure that my data is displayed 100% correctly through the code, I learned a lot. I learned that a lot of my time that I dedicate to doing work is after practice.  While logging this information, I realized that I am very tired while doing my homework.  Although I dedicate ample time to my studies, the quality is not as high as I would prefer due to the time of the day that I am working.
\newpage


# Reflection 

<!--
Write your reflection here in paragraph form. In particular, address...

* What difficulties in the data collection and analysis process did you encounter?  Identify two of your main hurdles in gathering accurate data.
* What implications do those difficulties have for future data collection and/or analysis projects if you were to repeat this project? 
* Did you have enough data? How much data do you think you'd need to collect in order to satisfactorily answer your question(s) of interest? Would it be hard to collect that data? Why or why not?
* As someone who provides data, what expectations do you have when you give your data (e.g. to Facebook, Google, MapMyRun, etc.)?  
* As someone who analyzes others' data, what ethical responsibilities do you have?

You may address these questions in any particular order. If it helps, you can make a subsection for each. 
-->

In gathering accurate data, it was difficult to be accurate because during my studies, I would take short five minute breaks.  Therefore the time is not 100% accurate.  Additionally, another issue is many of my days are structured the same.  Therefore, there is not a lot of variation in the data between days except for Fridays and the weekend.

In the future, I would like to test the time of day and number of tasks I complete.  I feel that I can make better use of my time in the morning to work when I am less tired. I would perhaps also like to repeat this project over the course of a full month and gather more intense data. 

I have an ethical responsibility to be truthful in order to accurately represent my time and dedications.  If I were to lie, then I would have less to learn about myself and my habits.



